Internal models of type theory and internal higher categories for the Univalent Foundations

### 2014-07-12 article-models-ttinternal.tex
progress on weakening algebras

\subsubsection{Currying for weakening}
Let $\cftalg{A}$ be a CFT-algebra in context $\Gamma$ and let
\begin{equation*}
\jhom
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
  {\cftfamalg{\cftfamalg{\cftalg{A}}}}
  {\cftwk{\cftalg{A}}}
\end{equation*}
be a CFT-homomorphism. The abstraction of the rules in \autoref{comp-ew} is the
condition that the diagram
\begin{equation}\label{eq:cftwk-cftextcurry}
\begin{tikzcd}[column sep=large]
\ctxwk
  {\ctxext{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}
  {\cftfamalg{\cftalg{A}}}
  \ar{r}{\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftwk{\cftalg{A}}}}
  \ar{dr}[swap]{\jcomp{}{\cftfamext}{\cftwk{\cftalg{A}}}}
& \ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftfamalg{\cftfamalg{\cftalg{A}}}}
  \ar{d}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
  \\
& \cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}
\end{tikzcd}
\end{equation}
of CFT-algebras in context 
$\ctxext
    {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
    {\cftalgf{\cftfamalg{\cftalg{A}}}}
$ %
commutes.

\begin{lem}
Let $\cftalg{A}$ be a CFT-algebra in context gamma with a CFT-homomorphism 
$ \jhom
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
  {\jcomp{}{\cftctxext}{\cftfamalg{\cftalg{A}}}}
  {\cftwk{\cftalg{A}}}
$
satisfying the judgmental equality of \autoref{eq:cftwk-cftextcurry}. Then we
have the judgmental equality
\begin{equation*}
\apply\cftwk{\cftalg{A}}{\gamma}{\apply\cftfamext{\gamma}{a}{p}}
\jdeq
\jcomp{}
    {\apply\cftwk{\cftalg{A}}{\gamma}{a}}
    {\apply\cftwk{\cftalg{A}}{\apply\cftctxext{\gamma}{a}}{p}}
\end{equation*}
of CFT-homomorphisms from $\subst{\gamma}{\cftfamalg{\cftalg{A}}}$ to 
$\subst{p}{{a}{{\gamma}{\cftalg{F_{F_{F_A}}}}}}$ in context $\Gamma$,
for every $\gamma:A$, $a:\subst{\gamma}{\cftalgf{\cftalg{A}}}$ and
$p:\subst{a}{{\gamma}{\cftalgf{\cftfamalg{\cftalg{A}}}}}$. 
\end{lem}

@@

>>>>

\subsubsection{Currying for weakening}
Let $\cftalg{A}$ be a CFT-algebra in context $\Gamma$ and let
\begin{equation*}
\jhom
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
  {\cftfamalg{\cftfamalg{\cftalg{A}}}}
  {\cftwk{\cftalg{A}}}
\end{equation*}
be a CFT-homomorphism. The abstraction of the rules in \autoref{comp-ew} is the
condition that the diagram
\begin{equation}\label{diag:cftwk-curry}
\begin{tikzcd}[column sep=large]
\ctxwk
  {\ctxext{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}
  {\cftfamalg{\cftalg{A}}}
  \ar{r}{\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftwk{\cftalg{A}}}}
  \ar{dr}[swap]{\jcomp{}{\cftfamext}{\cftwk{\cftalg{A}}}}
& \ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftfamalg{\cftfamalg{\cftalg{A}}}}
  \ar{d}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
  \\
& \cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}
\end{tikzcd}
\end{equation}
of CFT-algebras in context 
$\ctxext
    {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
    {\cftalgf{\cftfamalg{\cftalg{A}}}}
$ %
commutes.

\begin{lem}
Let $\cftalg{A}$ be a CFT-algebra in context gamma with a CFT-homomorphism 
$ \jhom
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
  {\jcomp{}{\cftctxext}{\cftfamalg{\cftalg{A}}}}
  {\cftwk{\cftalg{A}}}
$
satisfying the judgmental equality of \autoref{diag:cftwk-curry}. Then we
have the judgmental equality
\begin{equation*}
\apply\cftwk{\cftalg{A}}{\gamma}{\apply\cftfamext{\gamma}{a}{p}}
\jdeq
\jcomp{}
    {\apply\cftwk{\cftalg{A}}{\gamma}{a}}
    {\apply\cftwk{\cftalg{A}}{\apply\cftctxext{\gamma}{a}}{p}}
\end{equation*}
of CFT-homomorphisms from $\subst{\gamma}{\cftfamalg{\cftalg{A}}}$ to 
$\subst{p}{{a}{{\gamma}{\cftalg{F_{F_{F_A}}}}}}$ in context $\Gamma$,
for every $\gamma:A$, $a:\subst{\gamma}{\cftalgf{\cftalg{A}}}$ and
$p:\subst{a}{{\gamma}{\cftalgf{\cftfamalg{\cftalg{A}}}}}$. 
\end{lem}

@@

\begin{proof}
To show that $\cftwk{\cftfamalg{\cftalg{A}}}$ satisfies the Currying rule, we
must verify that the diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
\ctxwk
  {\ctxext{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
  {\cftfamalg{\cftfamalg{\cftalg{A}}}}
  \ar{r}{\ctxwk{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}{\cftwk{\cftfamalg{\cftalg{A}}}}}
  \ar{dr}[swap]{\jcomp{}{\jcomp{}{\cftctxext}{\cftfamext}}{\cftwk{\cftfamalg{\cftalg{A}}}}}
& \ctxwk{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}
  \ar{d}{\jcomp{}{\cftfamext}{\cftwk{\cftfamalg{\cftalg{A}}}}}
  \\
& \cftfamalg{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}
\end{tikzcd}
\end{equation*}
commutes judgmentally. We do this by showing that each of the ingredients of this
diagram is judgmentally equal to the corresponding ingredient of the diagram in
\autoref{diag:cftwk-curry} pulled back along $\cftctxext$. Thus, we are going
to show that
\begin{align*}
\ctxwk{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}{\cftwk{\cftfamalg{\cftalg{A}}}}
& \jdeq
\jcomp{}{\cftctxext}{\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftwk{\cftalg{A}}}}
  \tag{1}
  \\
\jcomp{}{\cftfamext}{\cftwk{\cftfamalg{\cftalg{A}}}}
& \jdeq
\jcomp{}{\cftctxext}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
  \tag{2}
  \\
\jcomp{}{\jcomp{}{\cftctxext}{\cftfamext}}{\cftwk{\cftfamalg{\cftalg{A}}}}
& \jdeq
\jcomp{}{\cftctxext}{\jcomp{}{\cftfamext}{\cftwk{\cftalg{A}}}}
  \tag{3}
\end{align*}
For the first, we have the judgmental equalities

To show that $\cftwk{\cftfamalg{\cftalg{A}}}$ preserves itself, we must verify
that the diagram
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
  \ar{d}
    [swap]{ \ctxwk
        {{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
        {\cftfamhom{\cftwk{\cftfamalg{\cftalg{A}}}}}
      }
  \ar{r}{\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\jcomp{}{\cftfamext}{\cftwk{\cftfamalg{\cftalg{A}}}}}}
& \ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftfamalg{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
  \ar{d}{\jcomp{}{\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\jcomp{}{\cftctxext}{\cftfamext}}}{\cftfamhom{\cftwk{\cftfamalg{\cftalg{A}}}}}}
  \\
\ctxwk{{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}{\jcomp{}{\cftwkc{\cftfamalg{\cftalg{A}}}}{\cftfamalg{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}}
  \ar{r}[swap,yshift=-1em]{%
    \jcomp{}
      { \cftwkf{\cftfamalg{\cftalg{A}}}}
      { \jcomp{}
          {\cftwkc{\cftfamalg{\cftalg{A}}}}
          { \jcomp{}
              {\cftfamext}
              { \jcomp{}
                  {\cftfamext}
                  {\cftwk{\cftfamalg{\cftalg{A}}}}
                }
            }
        }
    }
& \jcomp{}
    {\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\jcomp{}{\cftctxext}{\cftfamext}}}
    {\jcomp{}{\cftwkc{\cftfamalg{\cftalg{A}}}}{\cftfamalg{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}}
\end{tikzcd}
\end{equation*}
commutes judgmentally. We do this by showing that each of the ingredients of this
diagram is judgmentally equal to the corresponding ingredient of the diagram in
\autoref{diag:cftwk-cftwk} pulled back along $\cftctxext$. Thus, we are going
to show that
\begin{align*}
\ctxwk
  {\cftalgf{\cftfamalg{\cftalg{A}}}}
  {\jcomp{}{\cftfamext}{\cftwk{\cftfamalg{\cftalg{A}}}}}
& \jdeq
  \jcomp{}
    { \cftctxext}
    { \ctxwk{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
  \tag{1}
  \\
\jcomp{}
  {\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\jcomp{}{\cftctxext}{\cftfamext}}}
  {\cftfamhom{\cftwk{\cftfamalg{\cftalg{A}}}}}
& \jdeq
  \jcomp{}
    { \cftctxext}
    { \jcomp{}
        {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}
        {\cftfamhom{\cftwk{\cftalg{A}}}}
      }
  \tag{2}
  \\
\ctxwk
  {{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
  {\cftfamhom{\cftwk{\cftfamalg{\cftalg{A}}}}}
& \jdeq
  \jcomp{}
    { \cftctxext}
    { \ctxwk
        {{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}
        {\cftfamhom{\cftwk{\cftalg{A}}}}
      }
  \tag{3}
  \\
\jcomp{}
      { \cftwkf{\cftfamalg{\cftalg{A}}}}
      { \jcomp{}
          {\cftwkc{\cftfamalg{\cftalg{A}}}}
          { \jcomp{}
              {\cftfamext}
              { \jcomp{}
                  {\cftfamext}
                  {\cftwk{\cftfamalg{\cftalg{A}}}}
                }
            }
        }
& \jdeq
  \jcomp{}
    { \cftctxext}
    { \jcomp{}
        { \cftwkf{\cftalg{A}}}
        { \jcomp{}
            {\cftwkc{\cftalg{A}}}
            { \jcomp{}
                {\cftctxext}
                { \jcomp{}
                    {\cftctxext}
                    {\cftwk{\cftalg{A}}}
                  }
              }
          }
      }
  \tag{4}
\end{align*}
For the first, note that we have the judgmental equalities
\begin{align*}
\ctxwk
  {\cftalgf{\cftfamalg{\cftalg{A}}}}
  {\jcomp{}{\cftfamext}{\cftwk{\cftfamalg{\cftalg{A}}}}}
& \jdeq 
  \ctxwk
    {\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}
    {\jcomp{}{\cftfamext}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
  \\
& \jdeq
  \ctxwk
    {\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}
    {\jcomp{}{\cftctxext}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
  \\
& \jdeq
  \jcomp{}
    { \cftctxext}
    { \ctxwk{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
\end{align*}
For the second, note that we have the judgmental equalities
\begin{align*}
\jcomp{}
  {\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\jcomp{}{\cftctxext}{\cftfamext}}}
  {\cftfamhom{\cftwk{\cftfamalg{\cftalg{A}}}}}
& \jdeq
  \jcomp{}
    {\ctxwk{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}{\jcomp{}{\cftctxext}{\cftfamext}}}
    {\cftfamhom{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
  \\
& \jdeq
  \jcomp{}
    {\jcomp{}{\cftctxext}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}}
    {\cftfamhom{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
  \\
& \jdeq
  \jcomp{}
    {\jcomp{}{\cftctxext}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}}
    {\jcomp{}{\cftctxext}{\cftfamhom{\cftwk{\cftalg{A}}}}}
  \\
& \jdeq
  \jcomp{}
    { \cftctxext}
    { \jcomp{}
        {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}
        {\cftfamhom{\cftwk{\cftalg{A}}}}
      }
\end{align*}
For the third, note that we have the judgmental equalities
\begin{align*}
\ctxwk
  {{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
  {\cftfamhom{\cftwk{\cftfamalg{\cftalg{A}}}}}
& \jdeq
  \ctxwk
    {{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftfamalg{\cftalg{A}}}}}}
    {\cftfamhom{\cftwk{\cftfamalg{\cftalg{A}}}}}
  \\
& \jdeq
  \ctxwk
    {\jcomp{}{\cftctxext}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}}
    {\cftfamhom{\cftwk{\cftfamalg{\cftalg{A}}}}}
  \\
& \jdeq
  \ctxwk
    {\jcomp{}{\cftctxext}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}}
    {\cftfamhom{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
  \\
& \jdeq
  \ctxwk
    {\jcomp{}{\cftctxext}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}}
    {\jcomp{}{\cftctxext}{\cftfamhom{\cftwk{\cftalg{A}}}}}
  \\
& \jdeq
  \jcomp{}
    { \cftctxext}
    { \ctxwk
        {{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}
        {\cftfamhom{\cftwk{\cftalg{A}}}}
      }
\end{align*}
For the forth, note that we have the judgmental equalities
\begin{align*}
\jcomp{}
      { \cftwkf{\cftfamalg{\cftalg{A}}}}
      { \jcomp{}
          {\cftwkc{\cftfamalg{\cftalg{A}}}}
          { \jcomp{}
              {\cftfamext}
              { \jcomp{}
                  {\cftfamext}
                  {\cftwk{\cftfamalg{\cftalg{A}}}}
                }
            }
        }
& \jdeq
  \jcomp{}
      { \cftwkf{\cftfamalg{\cftalg{A}}}}
      { \jcomp{}
          { \jcomp{}{\cftctxext}{\cftwkc{\cftalg{A}}}}
          { \jcomp{}
              {\cftfamext}
              { \jcomp{}
                  {\cftfamext}
                  {\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
                }
            }
        }
  \\
& \jdeq
  \jcomp{}
      { \cftwkf{\cftfamalg{\cftalg{A}}}}
      { \jcomp{}
          { \jcomp{}{\cftctxext}{\cftwkc{\cftalg{A}}}}
          { \jcomp{}
              {\cftctxext}
              { \jcomp{}
                  {\cftctxext}
                  {\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
                }
            }
        }
  \\
& \jdeq
  \jcomp{}
      { \cftwkf{\cftfamalg{\cftalg{A}}}}
      { \jcomp{}
          { \cftctxext}
          { \jcomp{}
              { \cftwkc{\cftalg{A}}}
              { \jcomp{}
                  {\cftctxext}
                  {\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
                }
            }
        }
  \\
& \jdeq
  \jcomp{}
      { \jcomp{}{\cftctxext}{\cftwkf{\cftalg{A}}}}
      { \jcomp{}
          { \cftctxext}
          { \jcomp{}
              { \cftwkc{\cftalg{A}}}
              { \jcomp{}
                  {\cftctxext}
                  {\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
                }
            }
        }
  \\
& \jdeq
  \jcomp{}
    { \cftctxext}
    { \jcomp{}
        { \cftwkf{\cftalg{A}}}
        { \jcomp{}
            {\cftwkc{\cftalg{A}}}
            { \jcomp{}
                {\cftctxext}
                { \jcomp{}
                    {\cftctxext}
                    {\cftwk{\cftalg{A}}}
                  }
              }
          }
      }
\end{align*}
Thus we may conclude that the diagram above of which we had the aim to show that
it commutes judgmentally is a pullback of the diagram in \autoref{diag:cftwk-cftwk}
by $\cftctxext$ and therefore it commutes as desired.
\end{proof}
